---
title: "02 Data wrangling"
---

# 1) Setup  

```{r Setup, message=F}
# Loading packages (every time you open R)
library(readxl) # to read excel files
library(janitor) # to clean data
library(dplyr) # wrangling
library(tidyr) # wrangling
library(readr) # to export csv

```


```{r Reading data}
soyk <- read_excel("../data/01_soybeans.xlsx")

# Checking first rows
soyk

# View

```

This study was conducted at four different locations in one year to assess the effect of potassium (K) fertilizer rates on soybean trifoliate K concentration at growth stage R4.  

Each study was a randomized complete block design (RCBD) with K rate as the only treatment factor with levels 0, 50, 100, and 150 lbs K2O/ac.

```{r Glimpse}
glimpse(soyk)
```

```{r Summary}
summary(soyk)
```

# 2) Data wrangling  
```{r soyk_1 - clean_names()}
soyk_1 <- clean_names(soyk)

soyk_1
```


```{r soyk_2 - filter()}
names(soyk_1) 

soyk_2 <- filter(soyk_1,
                 k_rate_lbac == 0 & plot == 105
                 )

soyk_2
```

```{r soyk_3 - mutate()}
# Transforming K rate from lbs/ac to kg/ha
soyk_3  <- mutate(soyk_1,
                  k_rate_kgha = k_rate_lbac*0.454/0.405,
                  k_rate_kgha = round(k_rate_kgha, 0)
                  )

soyk_3
```

```{r soyk_4 - case_when()}
soyk_4 <- mutate(soyk_3,
                 rep = case_when(
                   plot < 200 ~ 1,
                   plot > 200 & plot < 300 ~ 2,
                  plot > 300 & plot < 400 ~ 3,
                   plot > 400 ~ 4
                 )
                 )

soyk_4

```

# pivot changes from wide to long format

```{r soyk_5 - pivot_longer()}
names(soyk_4)

soyk_5 <- pivot_longer(soyk_4,
                       names_to = "location.k",
                       values_to = "k_pct",
                       cols = contains("k_pct")
                       
                       )

soyk_5

View(soyk_5)
```

```{r soyk_6 - separate()}
soyk_6 <- separate(soyk_5,
                   location.k, # take this column and break down into three pieces
                   into = c("location", "k", "pct"),
                   sep = "_"
                   )

soyk_6

View(soyk_6)
```

```{r soyk_7 - select()}
names(soyk_6) # select = columns and filter = rows sorting

soyk_7 <- dplyr::select(soyk_6,
                 location, rep, k_rate_kgha, k_pct
                 )

soyk_7
```

```{r soyk_8 - arrange()}
soyk_8 <- arrange(soyk_7,
                  location
                  )

soyk_8
```

```{r soyk_9 and 10 - group_by and summarise}
# group_by
soyk_9 <- group_by(soyk_8,
                   location
                   
                   )

soyk_9

class(soyk_8)

class(soyk_9)

# summarise
soyk_10 <- summarise(soyk_9,
                     mean_pct = mean(k_pct)
                     )

soyk_10
```

```{r Piping it all together} 
soyk_w <- soyk %>% # Take the left function and assign its function to the right 
  clean_names() %>%
  mutate(
                  k_rate_kgha = k_rate_lbac*0.454/0.405,
                  k_rate_kgha = round(k_rate_kgha, 0)
                  ) %>%
  mutate(
                 rep = case_when(
                   plot < 200 ~ 1,
                   plot > 200 & plot < 300 ~ 2,
                  plot > 300 & plot < 400 ~ 3,
                   plot > 400 ~ 4
                 )
                 ) %>%
  pivot_longer(
                       names_to = "location.k",
                       values_to = "k_pct",
                       cols = contains("k_pct")
                       
                       )%>%
  
  separate(
                   location.k, # take this column and break down into three pieces
                   into = c("location", "k", "pct"),
                   sep = "_"
                   ) %>%
  
  dplyr::select(
                 location, rep, k_rate_kgha, k_pct
                 )  %>%

  arrange(
                  location
                  )


soyk_w


```

```{r Checking if same as without pipe}
# Same result?
identical(soyk_8, soyk_w)

```

```{r Saving to file for further analysis}
write_csv(soyk_w,
          "../output/soybean123.csv"
          )

```

# 3) Assignment   
Starting from soyk, use data wrangling functions in a pipe workflow to generate the average k_pct for each K rate (in kg/ha) within each location (basically, averaging over reps). You can use the chunk below to complete this quiz.

```{r assignment}
soy_average <- soyk %>% # Take the left function and assign its function to the right 
  clean_names() %>%
  mutate(
                  k_rate_kgha = k_rate_lbac*0.454/0.405,
                  k_rate_kgha = round(k_rate_kgha, 0)
                  ) %>%
  mutate(
                 rep = case_when(
                   plot < 200 ~ 1,
                   plot > 200 & plot < 300 ~ 2,
                  plot > 300 & plot < 400 ~ 3,
                   plot > 400 ~ 4
                 )
                 ) %>%
  pivot_longer(
                       names_to = "location.k",
                       values_to = "k_pct",
                       cols = contains("k_pct")
                       
                       )%>%
  
  separate(
                   location.k, # take this column and break down into three pieces
                   into = c("location", "k", "pct"),
                   sep = "_"
                   ) %>%
  
  dplyr::select(
                 location, rep, k_rate_kgha, k_pct
                 )  %>%

  arrange(
                  location,
                  
                  ) %>%

  group_by(
                   location,
                   k_rate_kgha
                   
                   ) %>%

  summarise(
                     avg_k_pct = mean(k_pct)
                     )
  
soy_average

```

Tip: the final data frame will have 16 rows.

Submit your result (take a screenshot that includes the code and the chunk output) on eLC under Assignment #2 by Monday Jan 26th 11:59 pm.



